---
title: "UC Berkeley Admissions Predictor"
author: "James Nacino"
date: "Monday, June 15, 2015"
output: html_document
---

##Question
Is it possible to predict admittance to a specific department at UC Berkeley based only on gender?

##Introduction
Getting into a prestigous school like UC Berkeley is tough. I would like to know if there is any correlation between admittance rate and the variables: gender and Department. Using the 'datasets' package in R, I have loaded the UCBAdmissions dataset. This dataset consists of three variables. 
```{r, echo=FALSE}
summary(UCBAdmissions)
head(data.frame(UCBAdmissions))
```

##Hypothesis
It is possible to predict a male or female student getting admitted into a specific department at UC Berkeley.

##Experiment
The department variable includes six different values (A,B,C,D,E, and F). From this we do not know if "A" corresponds to business, engineering, nursing, etc. I will collect all of the different departments of UC Berkeley through web scraping at their site "http://www.berkeley.edu/atoz/dept" and randomly select six different departments from the list. Then I will set the department values (A,B,C,D,E, and F) to one of the deparments that I randomly generated. Next, I will calculate the probability of each gender being accepted into the department. My shiny application will focus upon these probabilities. Each time the Shiny application will be loaded, there will be another set of random departments.

####Loading the packages and loading the dataset
```{r}
##Load packages
library(XML)
library(datasets)

##Create a data frame with the admissions data
df <- data.frame(UCBAdmissions)
head(df)
```

####Web scraping the UC Berkeley website to gather the list of their departments
```{r}
##Extract all departments from UC Berkeley webpages 
final_depts <- vector()
#Using 26 because there are 26 different webpages of departments. One page for each letter in the alphabet
for (i in 1:26) {
  letter <- LETTERS[i]
  deptsUrl <- "http://www.berkeley.edu/atoz/dept/"
  deptsUrl <- paste(deptsUrl,letter, sep="")
  doc <- htmlTreeParse(deptsUrl, useInternal=TRUE)
  depts <- xpathSApply(doc, "//div[@class='list-content']//a", xmlValue)
  depts <- sub("\n +", "", depts)
  final_depts <- append(final_depts, depts)
}
final_depts[1:40]
```

####Create the modified dataset replacing alphabet characters with actual department names
```{r}
##Pick six random departments at UCB
rand_depts_gen <- sample(final_depts, 6)

##Assign the values (A,B,C,D,E,F) of the departments column an actual department name from the randomly generate 
assign_depts <- vector()
for (i in 1:6){
temp <- rep(rand_depts_gen[i], times=4)
assign_depts <- append(assign_depts, temp)
}
df$steps <- assign_depts
```

Here is the modified dataset which replaces the departments with actual names.
```{r, echo=FALSE}
head(df)
```

####Modify the dataset again to include the probability of each gender being admitted into the department
```{r}
##Calculate the probabilities for gender admittance into each department
Probability <- vector()
for (i in 1:6){
  temp <- df[df$Dept==unique(df$Dept)[i],]
  temp_male <- temp[temp$Gender=="Male",]
  temp_male2 <- sum(temp_male$Freq)
  temp_male3 <- temp_male$Freq[1]/temp_male2
  Probability <- append(Probability, temp_male3)
  temp_male4 <- temp_male$Freq[2]/temp_male2
  Probability <- append(Probability, temp_male4)
  temp_fmale <- temp[temp$Gender=="Female",]
  temp_fmale2 <- sum(temp_fmale$Freq)
  temp_fmale3 <- temp_fmale$Freq[1]/temp_fmale2
  Probability <- append(Probability, temp_fmale3)
  temp_fmale4 <- temp_fmale$Freq[2]/temp_fmale2
  Probability <- append(Probability, temp_fmale4)
}
df <- cbind(df, Probability)
df$Probability <- round(df$Probability, digits=3) ##Round two three decimal places
df$Probability <- sprintf("%.1f %%", 100*df$Probability) ##Format probability as percentage
head(df)
```